---
title: "Choice-based conjoint analysis demonstration"
author: "Chris Chapman"
date: "`r Sys.Date()`"
# output: rmarkdown::html_vignette
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Choice-based conjoint analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(choicetools)
```


# Introduction to Choice-Based Conjoint Analysis

In this vignette, we demonstrate a simple choice-based conjoint (CBC) 
analysis project, end to end. 

Choice-based conjoint is a variety of conjoint analysis in which
respondents choose among several (typically 2-4) products that differ in 
various _attributes_ such as brand, performance, and price. Each attribute
will have two or more _levels_, such as brand name or specific prices. A 
complete product (aka 'concept') is a fully specified set of attributes 
and levels (Chapman & Feit, 2019).

For example, a Ford Truck with 300 HP engine at $35000 is an 
example of a product that 'conjoins' four attributes with specific levels: 
body style (with the level 'truck'), brand (Ford), engine size (300 HP), 
and price ($35000).

By randomizing the products' attributes and asking for repeated selection 
of preference, we can estimate the relative value (aka _utility_) of each
attribute level. For instance, we can determine whether the engine size is 
more important than brand. We can also determine willingness to pay. For 
example, we can match the utility scores for engine size to those for price, 
to see how much more we might expect respondents to pay for a specific engine
size.

Each respondent is asked to choose among a specific number of concepts
(products) at a time, such as their preference among 3 choices. They do this 
repeatedly, typically 8-12 times (aka _tasks_).

The CBC tools here are probably most useful for 
_didactic_ purposes --- to demonstrate and explain conjoint analysis in a 
classroom setting. A complete conjoint analysis study can be run with this
code and a simple spreadsheet editing program, such as Google Sheets or
Microsoft Office 365. (See notes at the end for details on that.) Beyond
that, many CBC studies with simple design characteristics can be estimated 
using the tools.


# Setup for a CBC study

In this vignette, we will imagine a hypothetical USB flash drive with five 
attributes: Brand (4 choices), Style (4 choices), Price (5 price points),
Color (2 choices), and Size (4 choices). Further, we assume that we will 
ask for preference among 3 concepts at a time, asking each respondent to 
judge 12 concepts (sets of concepts), with a total of N=400 respondents

We set up the study by defining those parameters:

```{r}
set.seed(98103)

cbc.tasks     <- 12   # trials per respondent
cbc.concepts  <- 3     # cards per trial
N             <- 400       # N of respondents

cbc.attrs     <- c(Brand=4, Style=4, Price=5, Color=2, Size=4)
```

Next, we can assign labels for the attribute levels. In the present code, 
this is a simple list, where you should ascertain that they are ordered 
correctly. We specify brands of "Alpha", "Bravo", and so forth; hypothetical  
design styles of "Shiny", "Flat", etc; and the other attributes as follow:

```{r}
cbc.levels    <- c("Alpha", "Bravo", "Charlie", "Delta",    # Brand
                   "Shiny", "Flat",  "Sparkly", "Odd",      # Style
                   "$9",  "$14",  "$19",  "$24",  "$29",    # Price
                   "Blue",  "Red",                          # Color
                   "8GB", "16GB", "32GB", "64GB")           # Size
```

We can create a product concept by randomly selecting one level for each of
the attributes. Thus, we might have an Alpha brand Shiny 16GB Blue USB drive 
for \$19, or a Charlie Flat design, Red 64GB device for \$9. Each task
in the survey will present 3 such randomly constructed concepts and ask the
respondent to select the single concept that is most preferred.

Now we create an experimental design matrix for the study. The 
design matrix specifies which randomly-generated concepts appear together.
To ensure the design is balanced and unbiased, we want each level to appear 
approximately the same number of times, and approximately the same number 
in combination with every other attribute. This is done with the 
`generalMNLrandomTab()` function:


```{r, echo=FALSE}
cbc.tab <- generateMNLrandomTab(cbc.attrs, respondents=N, 
                                cards=cbc.concepts, trials=cbc.tasks )
knitr::kable(head(cbc.tab))
```

This function iteratively searches randomized designs and retains a design if 
it has improved balance vs. the previous best design. The resulting design 
matrix specifies the level of each attribute that might be shown on the survey.

In most cases, you would obtain this design matrix from the survey
platform in which the survey was fielded. For instance, Sawtooth Software
can export the "tab design" matrix in a compatible format. In that case, you
do not need to generate a design matrix here; use that one instead. We 
generate the design here in order to field a survey for didactic purposes; 
see below.

We can also convert the "tab style" layout to a fully-extended, dummy coded 
matrix of 0 and 1 entries:

```{r, echo=FALSE, results='asis'}
cbc.des <- convertSSItoDesign(cbc.tab)              # dummy coded matrix
knitr::kable(head(cbc.des, 9))
```


In this case, we see that the first product concept is a combination of Brand 2,
Style 3, Price level 2, and so forth.

Given the design matrix, we can see how the survey choice tasks might 
appear (with minimal instruction). `writeCBCdesignCSV()` will produce a
minimal "survey" in CSV format:

```{r}
writeCBCdesignCSV(head(cbc.tab, 6), cards = 3, trials = 2, 
                  attr.list=cbc.attrs, lab.attrs=names(cbc.attrs),
                  lab.levels = cbc.levels)
```

By "fielding" such a survey in a spreadsheet editor --- such as Google 
Sheets or Microsoft Office --- we can easily demonstrate a CBC survey in 
a classroom setting. (To design a better looking and more robust survey 
for _real_ respondents, use a survey authoring platform such as Sawtooth
Software or Qualtrics.) We use the `digest` package (Eddelbuettel et al, 2018) 
to include a hash value for the design matrix as noted at the top of the CSV. 
This helps ensure that the collected data match the design matrix when choices 
are imported.


# Aggregate Logit Model Estimation

The simplest form of estimation is an aggregate multinomial logit model (MNL). This 
estimates a simple multinomial (choice among options) model for the entire sample. 
The estimated coefficients are known as *part worths* ("part" because the values
for multiple attributes are added together for the full utility of a concept). In 
an MNL model, the part worths sum to zero across the levels of each attribute.

To see how this works, we use `generateRNDpws()` to make up some zero-sum part worths, 
and then `pickMNLwinningCards()` to find the preferred concept based on those
part worths for each choice set:
```{r}
cbc.pws <- generateRNDpws(cbc.attrs)                # make up some zero-sum part worths

cbc.win <- pickMNLwinningCards(cbc.des, cbc.pws)          
knitr::kable(head(cbind(cbc.win, cbc.des), 9))
```

In the first set of choices -- the first three rows -- the second concept 
(Brand 1, Style 2, Price 3, etc.) was preferred and selected in the `cbc.win` 
vector. 

Now we estimate the part worths based on those "observed" choices, and plot
those estimates against the original part worths. We see a near perfect recovery 
of the part worth values (aside from minor rescaling of the magnitude):

```{r, fig.show='hold'}
cbc.mnl <- estimateMNLfromDesign(cbc.des, cbc.win, cards=cbc.concepts)
plot(cbc.pws, cbc.mnl)
```

# Write CSV Structure and Read Choices from It

Of course it is more interesting to collect data from real individuals! Most 
of the time one would do that by fielding a survey online using survey authoring
tools such as Sawtooth Software or Qualtrics. Such tools are able to 
display CBC tasks in a friendly way and require users to make tradeoffs among
concepts.

For this vignette, however, we will write out the CSV file, and then read 
the data back:

```{r}
csv.filename <- "~/Downloads/testCBC.csv"
writeCBCdesignCSV(cbc.tab, filename=csv.filename,   # filename="" for console
                  cards=3, trials=12, 
                  attr.list=cbc.attrs, lab.attrs=names(cbc.attrs),
                  lab.levels = cbc.levels, overwrite=TRUE)

# read the CSV 
csvfile.in  <- readLines(csv.filename)
```

If respondents had filled in choices, we could estimate the MNL model from them.
Because they have not, we will make some random choices and rewrite the CSV:
```{r}
# Fill in all of the choices with random choices (1, 2, or 3)
lines.with.choices <- which(grepl("CHOICE for Trial [0-9]+", csvfile.in))
csvfile.in[lines.with.choices] <- paste(csvfile.in[lines.with.choices], 
                                        sample(cbc.concepts, 
                                               length(lines.with.choices), 
                                               replace = TRUE))
writeLines(csvfile.in, con=csv.filename)
```
Now we have a CSV that might have been completed by respondents (if they
were answering randomly). We read those data using `readCBCchoices()` with
the design matrix (`cbc.tab`) and the filename. `readCBCchoices()` will check
that the design hash for the file matches the design matrix that we give it.
```{r}
# get those choices
cbc.choices <- readCBCchoices(cbc.tab, filename=csv.filename, 
                              cards=3, trials=12, verbose=FALSE)

```

We use `estimateMNLfromDesign()` to estimate the part worths from the observed
data. We plot those and see that -- because the choices were assigned randomly --
the estimates are mostly close to zero, and have no correspondence to the 
original part worths:
```{r}
# estimate partworths (should be near 0 because all choices were random)
cbc.mnl2 <- estimateMNLfromDesign(cbc.des, cbc.choices, cards=cbc.concepts,
                                  no.output = TRUE)
plot(cbc.pws, cbc.mnl2)
abline(h=0)
```

It is important to note that the `estimateMNLfromDesign()` function is not
intended for production usage; it is primarily a didactic function whose 
NR algorithm is easy to explain (and to read in code). For actual analyses,
we would use hierarchical Bayes estimation instead (Rossi, Allenby, and 
McCulloch, 2005).

# Hierarchical Bayes estimation

Hierachical Bayes estimation is typically used for choice models to estimate 
a mixed effects model. The *upper level* comprises fixed effects estimates
for the sample, while the *lower level* represents estimates for each 
individual respondent within the overall distribution. The individual-level 
estimates are based on the group of multiple observations for each respondent.

If you fielded the survey online, you would commonly receive responses as 1/0 
for each row, indicating whether the item was chosen (similar to our 
`cbc.win` vector above). For exposition here, we will use the winning cards
that were selected on the basis of our simulated part worths, given in `cbc.win`,
but add a degree of randomness.

To estimate the hierarchical Bayes model, use the function 
`estimateMNLfromDesignHB()`. This function is primarily a wrapper that 
simplifies the data setup and calls `ChoiceModelR::choicemodelr()` 
(Sermas, 2012). Note that we use the *tab* format for the design matrix 
(as opposed to the 1/0 dummy coded, expanded version). 

This is an iterative MCMC estimation routine. For speed, 
we specify 2000 total MCMC draws (burn-in and posterior); in practice, 
this would typically be 10000s, as the MCMC process will need a longer 
burn-in period for convergence. 
`estimateMNLfromDesignHB()` includes optional arguments to specify the 
proportion of the MCMC chain that is regarded as posterior 
draws  (`pitersUsed`), whether to save the draws for each respondent (`drawKeep`)
and the proportion of posterior draws to retain (skipping them in the sequence
to avoid autocorrelation; `drawKeepK`). 
```{r}
# replace 30% of the perfect "winning" vector with random draws
cbc.win2    <- cbc.win
cbc.replace <- sample(length(cbc.win2), length(cbc.win2)*0.3)  # replace these
cbc.win2[cbc.replace] <- sample(3, length(cbc.replace), replace=TRUE)

# estimate using the design and winning cards
cbc.hb <- estimateMNLfromDesignHB(tmp.des=cbc.tab, tmp.win=cbc.win2, 
                                  kCards=cbc.concepts, kTrials=cbc.tasks,
                                  kResp=N , mcmcIters=2000)
```

The MCMC process shows a trace plot of the estimates for each attribute. 
The estimated part worths may be
extracted from the HB model with the `extractHBbetas()` function. We extract those, 
add respondent IDs, and plot the values (Wickham, 2016; 
Wilke, 2018). We can plot each respondent
as a separate faint point, allowing us to see the distribution of estimates; 
this is especially helpful when there are groups of respondents who show differing
sets of preferences:

```{r}
cbc.est        <- data.frame(extractHBbetas(cbc.hb, cbc.attrs))
names(cbc.est) <- cbc.levels
cbc.est$ID     <- 1:nrow(cbc.est)

library(ggplot2)
library(reshape2)
cbc.m <- melt(cbc.est, id.vars = "ID")

library(ggridges)
ggplot(data=cbc.m, aes(x=value, y=variable, group=variable)) +
      geom_density_ridges(scale=0.9, alpha=0, jittered_points=TRUE,
                          rel_min_height=0.005,
                          position="points_sina",
                          point_color = "blue", point_alpha=1/sqrt(N),
                          point_size=2.5) +
        ylab("Item") + xlab("Relative preference (blue circles=individuals)") +
        ggtitle("Preference estimates: Distribution of individuals")

```

The `ggridges` plot is very useful to demonstrate the distribution of individuals
in HB estimates. If these were real data, we could interpret them as showing a 
strong preference for Sparkly design, Red color, $14 price point, and so forth.

Estimated share of preference (aka "market share") may be estimated using 
the `marketSim()` function. Suppose we wanted to assess preference for a 
Flat 8GB Blue drive at \$9 from Alpha, vs. an Odd 64GB Red drive at \$24 
from Bravo. We specify the sets of attributes for each product and compare 
them. In this case, we will use first choice preference, where
each individual is regarded as "purchasing" the item with highest preference; 
other options in `marketSim()` include share of preference at the individual
level, and first choice with randomization:

```{r}
# cbc.levels    <- c("Alpha", "Bravo", "Charlie", "Delta",    # Brand 1-4
#                    "Shiny", "Flat",  "Sparkly", "Odd",      # Style 5-8
#                    "$9",  "$14",  "$19",  "$24",  "$29",    # Price 9-13
#                    "Blue",  "Red",                          # Color 14-15
#                    "8GB", "16GB", "32GB", "64GB")           # Size 16-19

prod1 <- c(6, 16, 14,  9, 1)     # Flat 8GB Blue $9 Alpha
prod2 <- c(8, 19, 15, 12, 2)     # Odd 64GB Red $24 Bravo
usb.pref  <- marketSim(
  cbc.est,                    # matrix of individual-level utilities
  list(prod1, prod2),         # list of products to compare
  use.none=FALSE,             # we have no "none" column
  style="first")              # estimate share by first-choice approach

# see the overall preference share for the two products
colMeans(usb.pref)
```
Between just these two products, we estimate that 
`r round(colMeans(usb.pref)[1]*100)`% of respondents would prefer product 1, 
the $9 flat 8GB drive from Alpha.


# Classroom Usage


# Features Beyond CBC


# References


# References

Chapman, CN, and Feit, EMF (2019). *R for Marketing Research and Analytics, 2nd edition*. 
Chapter 13: Choice Modeling. New York: Springer.

Eddelbuettel, Dirk, with contributions by Antoine Lucas, Jarek Tuszynski, Henrik Bengtsson, Simon Urbanek,
  Mario Frasca, Bryan Lewis, Murray Stokely, Hannes Muehleisen, Duncan Murdoch, Jim Hester, Wush Wu, Qiang
  Kou, Thierry Onkelinx, Michel Lang, Viliam Simko, Kurt Hornik and Radford Neal. (2018). digest: Create
  Compact Hash Digests of R Objects. R package version 0.6.18. https://CRAN.R-project.org/package=digest
  
Rossi, PE, Allenby, GM, and McCulloch, RE (2005). *Bayesian Statistics and Marketing*. New York: Wiley.

Sermas, Ryan (2012). ChoiceModelR: Choice Modeling in R. R package version 1.2.
  https://CRAN.R-project.org/package=ChoiceModelR
  
Wickham, H. (2016). ggplot2: Elegant Graphics for Data Analysis. New York: Springer-Verlag.

Wilke, CO. (2018). ggridges: Ridgeline Plots in 'ggplot2'. R package version 0.5.1.
  https://CRAN.R-project.org/package=ggridges
